NumLin: Linear Types for Linear Algebra

### 2018-10-26 write-up/paper/intro.tex
Add points about finding bugs

\lang\ is a functional programming language designed to express the APIs of
low-level linear algebra libraries (such as BLAS/LAPACK) safely and explicitly.
It does so by combining linear types, fractional permissions, runtime errors
and recursion into a small, easily understandable, yet expressive set of core
constructs. In addition to this, \lang's implementation supports several
syntactic conveniences as well as a usable integration with real-world OCaml
code.

\subsection{Contributions}
\begin{itemize}
    \item We have designed the \lang\ programming language
    \item We illustrate that the design is sensible with many matrix-y examples
    \item We give a soundness proof for \lang, using a step-indexed logical relation
    \item Incredibly simple type inference algorihtm for polymorphic fractional permissions
        \begin{itemize}
            \item Compare to Bierhof et al's \emph{Fraction Polymorphic Permission Inference},
                which uses a fancy dataflow analysis
            \item We use exactly the same unification algorithm type polymorphism does
        \end{itemize}
    \item We have an implementation - compatible with and usable from existing code!
\end{itemize}

>>>>

\lang\ is a functional programming language designed to express the APIs of
low-level linear algebra libraries (such as BLAS/LAPACK) safely and explicitly.
It does so by combining linear types, fractional permissions, runtime errors
and recursion into a small, easily understandable, yet expressive set of core
constructs. \lang\ allows a novice to understand and work with complicated
linear algebra library APIs; as well as point out subtle aliasing bugs and
reduce memory usage in existing programs. In addition to this, \lang's
implementation supports several syntactic conveniences as well as a
\emph{usable} integration with real OCaml libraries.

\subsection{Contributions}

In this paper
\begin{itemize}
    \item we describe \lang, a linearly typed language for linear algebra programs
    \item we illustrate that \lang's design and features are well-suited to its
        intended domain with progressively sophisticated examples
    \item we prove \lang's soundness, using a step-indexed logical relation
    \item we describe a very simple, unification based type-inference algorithm
        for polymorphic fractional permissions (similar to ones used for
        parametric polymorphism), demonstrating an alternative approach to the
        dataflow analysis of Bierhof et al's \emph{Fraction Polymorphic
        Permission Inference}
    \item we describe an implementation that is both compatible with and usable
        from existing code.
\end{itemize}
